Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large-scale Public Dataset

### 2024-10-20 code/manuscript.Rmd
Changed site names and regrouped visualization + minor figure edits

---
title: "Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large-scale Public Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Durate ^1^
abstract: Conventional measurement and verification (M&V) methods for whole-building energy savings estimation are both time-consuming and unreliable, especially when non-routine events occur during the M&V process. Those events are unrelated to the proposed intervention strategy but have substaintial impacts on the building energy consumption. In this study, we argue that for switchable interventions (e.g. most of the control retrofits) can benefit from random sampling where the analyst randomly decide which strategy (i.e. baseline or intervention) to implement each day. We tested the novel randomized M&V method on a large public dataset which covers multiple climate zones and types of commercial buildings. We applied a virtual chilled water supply temperature reset based on outdoor weather as a control retrofit intervention. Our study shows that the new M&V method can estimate the savings accurately much quicker than the conventional method and most importantly, the estimation results are much more robust compared to the conventional method when non-routine events are present.

@@

# colors
ls_colors <- c("Baseline" = "#a6bddb",
               "Measured baseline" = "#a6bddb",
               "Adjusted baseline" = "#a6bddb",
               "Projected baseline\n(no change)" = "#2b8cbe",
               "Intervention" = "#fdbb84",
               "Measured interv" = "#fdbb84",
               "True savings" = "black",
               "True normalized savings" = "black")

@@

#### READ DATA ####
# Tidy dataset
readfile_tidy <- str_glue("../readfiles/tidy/")
fig_path = "../figs/manuscript/"

@@

```{r flowchart, fig.cap = "Workflow summary of the methodology of this paper", out.width="50%"}

@@

Therefor the resulting subset contains all buildings with 'tidy'
measurements. Figure \@ref(fig:tidysite) shows in total, the subset
contains `r nrow(all_names_tidy)` buildings in `r nrow(all_types_tidy)`
types from `r nrow(all_sites_tidy)` different climate zones.

```{r tidysite, fig.cap = "Site summary of the tidy building subset", fig.width=8, fig.height=6}

@@

p1 <- df_energy_tidy %>%
    group_by(type) %>%
    distinct(name) %>%
    summarise(n = n()) %>%
    ungroup() %>%
    mutate(type = fct_reorder(type, n, .desc = F)) %>%
    ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
    geom_col() +
    geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
    scale_y_continuous(expand = c(0, 0),
                       breaks = breaks_pretty(n = 4)) +
    scale_x_discrete(expand = c(0, 0.1)) +
    scale_fill_manual(values = type_colors) +
    labs(x = NULL,
         y = NULL,
         subtitle = "Across all sites",
         fill = NULL) +
    theme(axis.text.x = element_blank(),
          panel.grid.major.y = element_line(color = "grey80"),
          legend.direction = "horizontal",
          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  p2 <- df_energy_tidy %>%
    group_by(site, type) %>%
    distinct(name) %>%
    summarise(n = n()) %>%
    mutate(proportion = n / sum(n),
           total = sum(n),
           ymax = cumsum(proportion),
           ymin = c(0, head(ymax, n = -1))) %>%
    mutate(label_pos = (ymax + ymin) / 2) %>%
    ungroup() %>%
    group_by(type) %>%
    mutate(order = sum(n)) %>%
    ungroup() %>%
    mutate(type = fct_reorder(type, order, .desc = F)) %>%
    ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = type)) +
    geom_rect() +
    coord_polar(theta = "y") +
    xlim(c(2, 4)) +
    facet_wrap(~ site, nrow = 3) +
    labs(x = NULL,
         y = NULL,
         subtitle = "For each site",
         fill = NULL) +
    scale_fill_manual(values = type_colors) +
    geom_text(aes(x = 3.5, y = label_pos, label = n)) +
    geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black") +
    theme(legend.direction = "horizontal",
          axis.text = element_blank(),
          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  ggarrange(p1, p2,
            ncol = 2, nrow = 1,
            widths = c(0.2, 1),
            common.legend = TRUE,
            legend = "bottom") +
    plot_annotation(title = "Case study building type summary")
```

@@

Figure \@ref(fig:messysite) shows the summary of the 'messy' dataset,
which contains `r nrow(all_names_messy)` buildings in
`r nrow(all_types_messy)` types from `r nrow(all_sites_messy)` different
climate zones.

```{r messysite, fig.cap = "Site summary of the messy building subset", fig.width=8, fig.height=8}

p1 <- df_energy_messy %>%
    group_by(type) %>%
    distinct(name) %>%
    summarise(n = n()) %>%
    ungroup() %>%
    mutate(type = fct_reorder(type, n, .desc = F)) %>%
    ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
    geom_col() +
    geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
    scale_y_continuous(expand = c(0, 0),
                       breaks = breaks_pretty(n = 4)) +
    scale_x_discrete(expand = c(0, 0.1)) +
    scale_fill_manual(values = type_colors) +
    labs(x = NULL,
         y = NULL,
         subtitle = "Across all sites",
         fill = NULL) +
    theme(axis.text.x = element_blank(),
          panel.grid.major.y = element_line(color = "grey80"),
          legend.direction = "horizontal",
          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  p2 <- df_energy_messy %>%
    group_by(site, type) %>%
    distinct(name) %>%
    summarise(n = n()) %>%
    mutate(proportion = n / sum(n),
           total = sum(n),
           ymax = cumsum(proportion),
           ymin = c(0, head(ymax, n = -1))) %>%
    mutate(label_pos = (ymax + ymin) / 2) %>%
    ungroup() %>%
    group_by(type) %>%
    mutate(order = sum(n)) %>%
    ungroup() %>%
    mutate(type = fct_reorder(type, order, .desc = F)) %>%
    ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = type)) +
    geom_rect() +
    coord_polar(theta = "y") +
    xlim(c(2, 4)) +
    facet_wrap(~ site, nrow = 3) +
    labs(x = NULL,
         y = NULL,
         subtitle = "For each site",
         fill = NULL) +
    scale_fill_manual(values = type_colors) +
    geom_text(aes(x = 3.5, y = label_pos, label = n)) +
    geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black") +
    theme(legend.direction = "horizontal",
          axis.text = element_blank(),
          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  ggarrange(p1, p2,
            ncol = 2, nrow = 1,
            widths = c(0.2, 1),
            common.legend = TRUE,
            legend = "bottom") +
    plot_annotation(title = "Case study building type summary")
```

@@

```{r chwst, fig.cap = "Proposed chilled water supply temperature reset based on outdoor temperature", fig.width=8, fig.height=4}

data.frame(
  OAT = c(10, 15, 25, 30),  
  Baseline = c(6, 6, 6, 6),      
  Intervention = c(12, 12, 7, 7)
  ) %>% 
  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
  ggplot(aes(x = OAT, y = value, color = strategy)) +
  geom_line() +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = c(4, 6, 8, 10, 12),
                     limits = c(5.5, 12.5),
                     labels = number_format(suffix = " °C")) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(10, 15, 20, 25, 30), 
                     labels = number_format(suffix = " °C")) +
  labs(x = "Outdoor temperature", 
       y = "Chilled water supply temperature", 
       title = "Chilled water supply temperature reset on outdoor temperature") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

|                        |                                                                  |       |
|------------------|------------------------------------|------------------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

Figure \@ref(fig:timeline) shows the average estimated timeline for each
building across all sites and climate conditions if using randomized M&V
for the applied chilled water supply temperature reset intervention. The
timeline figures for each individual building are attached in the
supplementary material. The figure shows for all climate zones, covering
a sufficient range of outdoor weather condition is the most stringent
requirement. However, most buildings can achieve 80% of the required
range within 6 months, and all buildings can determine the savings,
including associated uncertainties, within 9 months---significantly
shorter than the baseline measurement period required by the
conventional method.

```{r timeline, fig.cap="Average randomized M&V timeline summary for all buildings at each site", fig.width=8, fig.height=4}

df_sprt_all_messy %>% 
  filter(seq == "eob" | seq == "temp") %>% 
  bind_rows(df_sprt_all_tidy %>% filter(seq == "eob" | seq == "temp")) %>% 
  group_by(site, seq) %>% 
  summarise(n_weeks = mean(n_weeks)) %>% 
  ungroup() %>% 
  mutate(seq = as.factor(seq), 
         seq = recode_factor(seq, "temp" = "80% temperature range satisfied", "eob" = "remaining weeks till finish")) %>% 
  ggplot() +
  geom_col(aes(x = site, y = n_weeks, fill = seq), position = "identity") +
  labs(x = "Site", 
       y = "Number of weeks", 
       fill = NULL, 
       title = "Average finishing time for all buildings using randomized M&V") +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = seq(0, 36, by = 12)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

```{r norm, fig.cap="Comparison of normalized annual fractional savings on site TMY between conventional and randomized method", fig.width=8, fig.height=4}

p1 <- df_FS_tidy %>% 
  bind_rows(df_FS_messy) %>% 
  filter(method == "true") %>% 
  mutate(method = "true savings") %>% 
  ggplot(aes(x = method, y = savings, fill = method)) +
  geom_errorbar(aes(ymin = min(savings), ymax = max(savings)), width = 0.2) +
  geom_boxplot(outlier.size = 2) +
  geom_text(aes(x = 1.35, 
                y = median(savings) + 1.5, 
                label = paste0("median: ", round(median(savings), digits = 0), "%")), 
            check_overlap = T) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     limits = c(0, 30), 
                     labels = number_format(suffix = "%")) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

p2 <- df_FS_tidy %>% 
  bind_rows(df_FS_messy) %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  left_join(df_seq_FS_tidy %>% 
              bind_rows(df_seq_FS_messy) %>% 
              filter(seq == "eob"), by = c("name", "site")) %>% 
  pivot_longer(c(conv, rand, FS), names_to = "parameter", values_to = "value") %>% 
  filter(parameter != "rand") %>% 
  mutate(parameter = as.factor(parameter), 
         parameter = recode_factor(parameter, 
                                   "conv" = "Conventional M&V", 
                                   "FS" = "Randomized M&V")) %>% 
  ggplot(aes(x = parameter, y = value, fill = parameter)) +
  stat_boxplot(geom ='errorbar', width = 0.5) +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     limits = c(0, 30)) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        axis.text.y = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  ggarrange(p1, p2,
            labels = c("a)", "b)"),
            ncol = 2, nrow = 1,
            widths = c(0.5, 1),
            legend = "none") +
    plot_annotation(title = "Normalized annual fractional savings estimation comparison", 
                    subtitle = "between convention and randomized M&V (with early stop)")
```

@@

```{r tidyfrac, fig.cap="Absolute deviation of M&V estimated fractional savings from true target savings using tidy subset", fig.width=8, fig.height=4}

@@

p1 <- plot_data %>% 
  arrange(abs_diff) %>% 
  ggplot() +
  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff), 
           stat = "identity", alpha = 0.5) +
  geom_hline(yintercept = median_rand, 
             linetype = "dashed", 
             color = "red") +
  annotate(y = median_rand + 0.5, 
           x = 20, 
           geom = "text", 
           label = paste0("(", "50% buildings: < ", round(median_rand, digits = 1), " %)"),
           size = 4) +
  geom_hline(yintercept = upper_rand, 
             linetype = "dashed", 
             color = "red") +
  annotate(y = upper_rand + 0.5, 
           x = 20, 
           geom = "text", 
           label = paste0("(", "95% buildings: < ", round(upper_rand, digits = 1), " %)"),
           size = 4) +
  scale_x_continuous(expand = c(0.02, 0), 
                     breaks = breaks_pretty(n = 3)) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     labels = number_format(suffix = " %"),
                     limits = c(0, max(plot_data$plot_max) + 0.5)) +
  labs(x = "Number of buildings", 
       y = "Absolute difference in measured savings", 
       subtitle = "Randomized M&V at early stop") +
  coord_cartesian(ylim = c(0, 10)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

```{r messyfrac, fig.cap="Absolute deviation of M&V estimated fractional savings from true target savings using messy subset", fig.width=8, fig.height=4}

@@

In this section, we demonstrate the influence of non-routine events
through two scenarios. In the first scenario, we quantified how much
occupancy would influence the electricity usage and added to the tidy
measurement set as they originally indicates no such change. In the
second scenario, we simply run the two M&V methods on the messy
measurement sets before adding the chilled water supply temperature
reset. Since no intervention is added, the more reliable method should
detect a 'saving' closer to 0.

@@

change_start <- occ_params$change_start
change_end <- occ_params$change_end
change <- occ_params$change

base_pre_meas <- df_base_conv %>%
  filter(datetime < as.Date("2017-01-01")) %>%
  mutate(eload = ifelse(month(datetime) >= month(change_start) & month(datetime) <= month(change_end), eload * (1 - change / 100), eload))

@@

```{r occchange, fig.cap="Occupancy change impact on TOWT model fitting and baseline projection", fig.width=8, fig.height=4}

>>>>

---
title: "Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large-scale Public Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Durate ^1^
abstract: Conventional measurement and verification (M&V) methods relying on pre- and post-  retrofit comparison for estimating whole-building energy savings are often time-consuming and unreliable, especially when non-routine events such as reduced occupancy occur during the M&V process. Those events are unrelated to the intervention but significantgly affect building energy consumption and thus when the analyst applies conventiona M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most of the control retrofits, can benefit from random sampling, where the analyst randomly selects whether to implement the baseline or the intervention strategy each day. We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that comparing to the conventional method, the randomized method provides faster and more accurate savings estimation. Additionaly we found that the randomized M&V approach estimates much closer to the true savings when non-routine events are present demonstrating much improved reliability.

@@

# colors
ls_colors <- c("Baseline" = "#c6dbef",
               "Measured baseline" = "#c6dbef",
               "Adjusted baseline" = "#c6dbef",
               "Projected baseline" = "#2171b5",
               "Intervention" = "#fdbb84",
               "Measured interv" = "#fdbb84",
               "Conventional M&V" = "grey70", 
               "Randomized M&V" = "#66c2a4", 
               "Buildings satisfying all criteria" = "#66c2a4", 
               "Buildings satisfying 80% TMY range" = "black", 
               "Buildings satisfying SPRT" = "black")

@@

#### READ DATA ####
df_loc <- read.csv("../readfiles/loc_map.csv")

# Tidy dataset
readfile_tidy <- str_glue("../readfiles/tidy/")
fig_path = "../figs/manuscript/"

@@

```{r flowchart, fig.cap = "Flow chat showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="50%"}

@@

Therefor the resulting subset contains all buildings with 'tidy'
measurements. Figure \@ref(fig:tidysite) shows in total, the subset
contains `r nrow(all_names_tidy)` buildings of `r nrow(all_types_tidy)`
types.

```{r tidysite, fig.cap = "Site summary of the tidy building subset (counts < 2 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=6}

@@

p1 <- df_energy_tidy %>%
  group_by(type) %>%
  distinct(name) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(type = fct_reorder(type, n, .desc = F)) %>%
  ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
  geom_col() +
  geom_text(aes(label = ifelse(n > 2, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = breaks_pretty(n = 4)) +
  scale_x_discrete(expand = c(0, 0.1)) +
  scale_fill_manual(values = type_colors) +
  labs(x = NULL,
       y = NULL,
       subtitle = "Combining all locations",
       fill = NULL) +
  theme(axis.text.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey80"),
        legend.direction = "horizontal",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  p2 <- df_energy_tidy %>%
    left_join(df_loc, by = "site") %>% 
    group_by(location, type) %>%
    distinct(name) %>%
    summarise(n = n()) %>%
    mutate(proportion = n / sum(n),
           total = sum(n),
           ymax = cumsum(proportion),
           ymin = c(0, head(ymax, n = -1))) %>%
    mutate(label_pos = (ymax + ymin) / 2) %>%
    ungroup() %>%
    group_by(type) %>%
    mutate(order = sum(n)) %>%
    ungroup() %>%
    mutate(type = fct_reorder(type, order, .desc = F)) %>%
    ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = type)) +
    geom_rect() +
    coord_polar(theta = "y") +
    xlim(c(2, 4)) +
    facet_wrap(~ location, nrow = 3) +
    labs(x = NULL,
         y = NULL,
         subtitle = "For each location",
         fill = NULL) +
    scale_fill_manual(values = type_colors) +
    geom_text(aes(x = 3.5, y = label_pos, label = ifelse(n > 2, as.character(n), ""))) +
    geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black") +
    theme(legend.direction = "horizontal",
          axis.text = element_blank(),
          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  ggarrange(p1, p2,
            ncol = 2, nrow = 1,
            widths = c(0.2, 1),
            common.legend = TRUE,
            legend = "bottom") +
    plot_annotation(title = "Summary of building numbers and types for 'tidy' subset")
```

@@

Figure \@ref(fig:messysite) shows the summary of the 'messy' dataset,
which contains `r nrow(all_names_messy)` buildings in
`r nrow(all_types_messy)` types.

```{r messysite, fig.cap = "Site summary of the messy building subset (counts < 2 are omitted for visualization; left: aggragated counts of buildings for each type; right: breakdown building counts for each building type at each location)", fig.width=8, fig.height=8}

p1 <- df_energy_messy %>%
    group_by(type) %>%
    distinct(name) %>%
    summarise(n = n()) %>%
    ungroup() %>%
    mutate(type = fct_reorder(type, n, .desc = F)) %>%
    ggplot(aes(x = 1, y = n, fill = as.factor(type))) +
    geom_col() +
    geom_text(aes(label = ifelse(n > 10, as.character(n), "")), color = "black", position = position_stack(vjust = 0.5)) +
    scale_y_continuous(expand = c(0, 0),
                       breaks = breaks_pretty(n = 4)) +
    scale_x_discrete(expand = c(0, 0.1)) +
    scale_fill_manual(values = type_colors) +
    labs(x = NULL,
         y = NULL,
         subtitle = "Combining all locations",
         fill = NULL) +
    theme(axis.text.x = element_blank(),
          panel.grid.major.y = element_line(color = "grey80"),
          legend.direction = "horizontal",
          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  p2 <- df_energy_messy %>%
    left_join(df_loc, by = "site") %>% 
    group_by(location, type) %>%
    distinct(name) %>%
    summarise(n = n()) %>%
    mutate(proportion = n / sum(n),
           total = sum(n),
           ymax = cumsum(proportion),
           ymin = c(0, head(ymax, n = -1))) %>%
    mutate(label_pos = (ymax + ymin) / 2) %>%
    ungroup() %>%
    group_by(type) %>%
    mutate(order = sum(n)) %>%
    ungroup() %>%
    mutate(type = fct_reorder(type, order, .desc = F)) %>%
    ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = type)) +
    geom_rect() +
    coord_polar(theta = "y") +
    xlim(c(2, 4)) +
    facet_wrap(~ location, nrow = 3) +
    labs(x = NULL,
         y = NULL,
         subtitle = "For each location",
         fill = NULL) +
    scale_fill_manual(values = type_colors) +
    geom_text(aes(x = 3.5, y = label_pos, label = ifelse(n > 1, as.character(n), ""))) +
    geom_text(aes(x = 2, y = 0, label = paste0("Total\n", total)), color = "black") +
    theme(legend.direction = "horizontal",
          axis.text = element_blank(),
          plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  ggarrange(p1, p2,
            ncol = 2, nrow = 1,
            widths = c(0.2, 1),
            common.legend = TRUE,
            legend = "bottom") +
    plot_annotation(title = "Summary of building numbers and types for 'messy' subset")
```

@@

```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature", fig.width=8, fig.height=4}

data.frame(
  OAT = c(10, 15, 25, 30),  
  Baseline = c(6, 6, 6, 6),      
  Intervention = c(12, 12, 7, 7)
  ) %>% 
  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
  ggplot(aes(x = OAT, y = value, color = strategy)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = c(4, 6, 8, 10, 12),
                     limits = c(5.5, 12.5),
                     labels = number_format(suffix = " °C")) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(10, 15, 20, 25, 30), 
                     labels = number_format(suffix = " °C")) +
  labs(x = "Outdoor temperature", 
       y = "Chilled water supply temperature", 
       color = NULL, 
       title = "Intervention logic: chilled water supply temperature reset") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

|                        |                                                                  |       |
|-------------------|----------------------------------|-------------------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

Figure \@ref(fig:timeline) shows the average estimated timeline for each
building across all sites and climate conditions if using randomized M&V
for the applied chilled water supply temperature reset intervention. The
timeline figures for each individual building are attached in the
supplementary material. The figure shows all buildings can detect a
saving statistically from sampled measurements within 12 months, which
is equivalent to the minimum test criterion. Similar to the findings
from our pervious study, covering a sufficient range of outdoor weather
condition is the most stringent requirement. If target buildings are
located in a climate zone with more preferable outdoor weather
conditions such as California, covering 80% of the TMY range should only
take 3 \~ 4 months. In this case, the analyst can conclude the M&V
simultaneously if the blocking period ends at the same time. However,
given most buildings requires 6 months to achieve 80% of the TMY range,
all buildings can determine the savings, including associated
uncertainties, within 9 months---significantly shorter than the baseline
measurement period required by the conventional method.

```{r timeline, fig.cap="Randomized M&V timeline of satisfying key stopping criteria for all buildings", fig.width=8, fig.height=4}

df_time <- df_sprt_all_messy %>% 
  filter(seq != "final") %>% 
  bind_rows(df_sprt_all_tidy %>% filter(seq != "final")) %>% 
  select(name, seq, n_weeks)

count <- list()

for (i in seq(1, 36, by = 1)){

  df_sprt <- df_time %>% 
    filter(seq == "sprt",
           n_weeks <= i)

  df_eob <- df_time %>% 
    filter(seq == "eob",
           n_weeks <= i)

  df_temp <- df_time %>% 
    filter(seq == "temp", 
           n_weeks <= i)

  count[[i]] <- tibble("n_weeks" = i, 
                       "eob" = nrow(df_eob), 
                       "temp" = nrow(df_temp), 
                       "sprt" = nrow(df_sprt))
}

count <- bind_rows(count)

count %>% 
  ggplot() +
  geom_bar(data = . %>% filter(n_weeks == 24 | n_weeks == 36), 
           aes(x = n_weeks, y = eob, fill = "Buildings satisfying all criteria"), 
           stat = "identity", 
           alpha = 0.8, 
           width = 1) +
  geom_line(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = temp, color = "Buildings satisfying 80% TMY range"), 
             size = 1.5) +
  geom_line(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"),
            alpha = 0.2) +
  geom_point(aes(x = n_weeks, y = sprt, color = "Buildings satisfying SPRT"), 
             size = 1.5, 
             shape = 17) +
  geom_vline(xintercept = seq(12, 36, by = 12), lty = "dashed", color = "grey80") +
  annotate(geom = "text", 
           x = seq(3, 33, by = 12), 
           y = 700, 
           label = paste0("12-week block"), 
           alpha = 0.5) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = ls_colors) +
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 38),
                     breaks = c(12, 24, 36), 
                     labels = c("12\nweeks", "24\nweeks", "36\nweeks")) +
  coord_cartesian(ylim = c(0, 720)) +
  labs(x = NULL, 
       y = "Number of buildings", 
       fill = NULL, 
       color = NULL, 
       title = "Overall timeline for buildings using randomized M&V") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

```{r norm, fig.cap="Comparison of annual fractional savings normalized on TMY weather conditions between conventional and randomized M&V method", fig.width=8, fig.height=4}

p1 <- df_FS_tidy %>% 
  bind_rows(df_FS_messy) %>% 
  filter(method == "true") %>% 
  mutate(method = "True savings") %>% 
  ggplot(aes(x = method, y = savings, fill = method)) +
  geom_errorbar(aes(ymin = min(savings), ymax = max(savings)), width = 0.2) +
  geom_boxplot(outlier.size = 2) +
  geom_text(aes(x = 1.35, 
                y = median(savings) + 1.5, 
                label = paste0("median: ", round(median(savings), digits = 0), "%")), 
            check_overlap = T) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     limits = c(0, 28), 
                     labels = number_format(suffix = "%")) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

p2 <- df_FS_tidy %>% 
  bind_rows(df_FS_messy) %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  left_join(df_seq_FS_tidy %>% 
              bind_rows(df_seq_FS_messy) %>% 
              filter(seq == "eob"), by = c("name", "site")) %>% 
  pivot_longer(c(conv, rand, FS), names_to = "parameter", values_to = "value") %>% 
  filter(parameter != "rand") %>% 
  mutate(parameter = as.factor(parameter), 
         parameter = recode_factor(parameter, 
                                   "conv" = "Conventional M&V", 
                                   "FS" = "Randomized M&V")) %>% 
  ggplot(aes(x = parameter, y = value, fill = parameter)) +
  stat_boxplot(geom ='errorbar', width = 0.5) +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     limits = c(0, 28)) +
  scale_fill_manual(values = ls_colors) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        axis.text.y = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  ggarrange(p1, p2,
            labels = c("a)", "b)"),
            ncol = 2, nrow = 1,
            widths = c(0.5, 1),
            legend = "none") +
    plot_annotation(title = "Normalized annual fractional savings estimation comparison")
```

@@

```{r tidyfrac, fig.cap="Absolute deviation of fractional savings from true target savings shown for the 'tidy' subset", fig.width=8, fig.height=4}

@@

p1 <- plot_data %>% 
  arrange(abs_diff) %>% 
  ggplot() +
  geom_bar(aes(x = seq(1, nrow(all_names_tidy), by = 1), y = abs_diff, fill = "Randomized M&V"), 
           stat = "identity", alpha = 0.5) +
  geom_hline(yintercept = median_rand, 
             linetype = "dashed", 
             color = "black") +
  annotate(y = median_rand + 0.5, 
           x = 20, 
           geom = "text", 
           label = paste0("(", "50% buildings: < ", round(median_rand, digits = 1), " %)"),
           size = 4) +
  geom_hline(yintercept = upper_rand, 
             linetype = "dashed", 
             color = "black") +
  annotate(y = upper_rand + 0.5, 
           x = 20, 
           geom = "text", 
           label = paste0("(", "95% buildings: < ", round(upper_rand, digits = 1), " %)"),
           size = 4) +
  scale_x_continuous(expand = c(0.02, 0), 
                     breaks = breaks_pretty(n = 3)) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     labels = number_format(suffix = " %"),
                     limits = c(0, max(plot_data$plot_max) + 0.5)) +
  scale_fill_manual(values = ls_colors) +
  labs(fill = NULL, 
       x = "Number of buildings", 
       y = "Absolute difference of fractional savings", 
       subtitle = "Randomized M&V") +
  coord_cartesian(ylim = c(0, 10)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

```{r messyfrac, fig.cap="Absolute deviation of fractional savings from true target savings shown for the 'messy' subset", fig.width=8, fig.height=4}

@@

We included two examples in this section to illustrate the impact of
non-routine events on M&V savings estimation. In the first scenario, we
manually subtracted a portion of the electricity usage for a specific
period in the tidy subset to simulate occupancy-confounded measurements,
which were not originally present in the dataset. In the second
scenario, focusing on the messy subset, which already contains
underlying changes, we removed the intervention effect and applied both
M&V methods to assess which method could more accurately detect the
resulting zero 'savings.'

@@

change_start <- occ_params$change_start
change_end <- occ_params$change_end
change <- mean(df_base_conv$eload) * occ_params$change / 100

base_pre_meas <- df_base_conv %>%
  filter(datetime < as.Date("2017-01-01")) %>%
  mutate(eload = ifelse(month(datetime) >= month(change_start) & month(datetime) <= month(change_end), eload - change, eload))

@@

```{r occchange, fig.cap="Illustration of occupancy change impact on TOWT model fitting and baseline projection in the post-retrofit period", fig.width=9, fig.height=4}
